---
title: "spatial_autocorrelation"
author: "Aimara Planillo"
date: '2022-10-04'
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of spatial autocorrelation in roadkill data

```{r packages}
library(tidyverse)
library(sf)
library(tmap)
library(dplyr)
library(purrr)
library(spatial)
library(spdep)
library(ggplot2)
library(MASS)
library(nlme)
library(ggeffects)
library(glmmTMB)
```

# Load and prepare data
```{r}
data_roadkills <- read_csv(paste0("data/GLM/table_glm_10.csv")) %>%
  mutate(road_importance = factor(road_importance, levels = c("minor_road", "major_road")),
         woodland2 = woodland^2,
         distance_amenity_km2 = distance_amenity_km^2)

str(data_roadkills)
head(data_roadkills)

## make spatial: all middle points of road segments
roadkills_sf <- st_as_sf(data_roadkills, coords = c("long_midpoint", "lat_midpoint"),
                         crs = 4326)

# Explore number of carcasses
tmap_mode("view")
tm_shape(roadkills_sf) +
  # tm_dots("red") + # this plots just the middle point of the road segments
  tm_dots("nbr_carcasses")
```



# 1- Spatial autocorrelation (SAC)

Variogram and correlogram from package "spatial"
For this we create a "trend surface" from spatial data and response variable.
We use nint = number of bins or intervals to use for checking for the correlation; they need to have 6 pairs of points or more.
To run Moran's I we use package "spdep"
This library uses the coordinates in a matrix and a "weights list" objects, class lw,
obtained from a "neighbour" object, class nb, with the function nb2listw
The neighbour object is a list of all localities. Each element is a vector of integers
that represents the identity of the localities that share borders with the target locality
(information about the neighbours of each point).
We use 4 neighbours for this (k = 4).

## Check for SAC in point data

```{r}
## project the coordinates to have the distances in meters
roadkills_sf_proj <- st_transform(roadkills_sf, crs = 32636)

tm_shape(roadkills_sf_proj) +
  tm_dots("nbr_carcasses")
```

```{r}
## compute moran test
moran_roadills_df <- roadkills_sf_proj %>%
  mutate(x_coord = st_coordinates(roadkills_sf_proj)[,1],
         y_coord = st_coordinates(roadkills_sf_proj)[,2]) %>% 
  st_drop_geometry() %>% 
  dplyr::select(x_coord,y_coord,nbr_carcasses) 

mycoord <- matrix(cbind(moran_roadills_df$x_coord,moran_roadills_df$y_coord), ncol=2)
mycoord.nb <- knn2nb(knearneigh(mycoord, k = 4))
mycoord.lw <- nb2listw(mycoord.nb, style="W")
plot(mycoord.nb,mycoord) # Represents each point linked to the nearest 4 neighbours 

roadkills_moran <- moran.test(moran_roadills_df$nbr_carcasses, mycoord.lw)

roadkills_moran$estimate
roadkills_moran$p.value
# moran test is significant: spatial autocorrelation!

## plot correlogram
dts2 <- list()
surface2 <- list()

dts2 <- data.frame(x = moran_roadills_df$x_coord, 
                   y = moran_roadills_df$y_coord, 
                   z = moran_roadills_df$nbr_carcasses)
surface2 <- surf.ls(2, dts2)
png("Correlogram_roadkill_segments.png", width = 600, height = 550)
correlogram(krig = surface2, nint = 100, plotit=TRUE, xlim=c(0,20000),
            xlab="distance (m)", ylab="Moran's I",
            cex.lab=1.2, cex.axis=1.2, cex.main = 1.2,
            main = "Correlogram - roadkill data segments")
abline(v= 5000, col = "blue", lty = 4)
dev.off()
```

There is correlation until 5 km

```{r}
surface2 <- list()
dts2 <- data.frame(x = moran_roadills_df$x_coord, 
                   y = moran_roadills_df$y_coord, 
                   z = moran_roadills_df$nbr_carcasses)
surface2 <- surf.ls(2, dts2)
x <- correlogram(krig = surface2, nint = 100, plotit = F, xlim=c(0,20000), 
            xlab="distance (m)", ylab="Moran's I",
            cex.lab=1.2, cex.axis=1.2, cex.main = 1.2, 
            main = "Correlogram - roadkill data segments")
data_plot_correlogram <- data.frame(x = x$x/1000,
                                    y = x$y)

range(data_plot_correlogram$y[18:97])

ggplot(data = data_plot_correlogram, aes(x = x, y = y)) +
  geom_point() + #shape = 1) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 5, linetype = "dotted", color = "blue") +
  xlim(c(0, 20)) +
  ylim(c(-1, 1)) +
  theme_classic() +
  labs(x = "distance (km)", y = "Moran's I")

ggsave("figure S2 (correlogram).png", width = 14, height = 10, units = "cm")
```


# 2- Select spatial structure

We are going to "cheat" and use null models with gls to do this because the full model is more difficult to evaluate

```{r}
## prepare data with projected coordinates
nrow(data_roadkills)
nrow(moran_roadills_df)

roadkills_df <- data_roadkills %>% 
  mutate(x_coord = st_coordinates(roadkills_sf_proj)[,1],
         y_coord = st_coordinates(roadkills_sf_proj)[,2]) %>% 
  mutate(group = 1)
  
head(roadkills_df)
str(roadkills_df)
```


```{r}
mod_formula <- formula("nbr_carcasses ~ 1")

# no spatial structure
fm0 <- gls(mod_formula, 
           data = roadkills_df,
           # correlation = corGaus(form=~x_coord+y_coord, nugget = T)
           )

# continous autoregressive process
fm_CAR <- gls(mod_formula, 
              data = roadkills_df,
              correlation = corCAR1(form=~x_coord+y_coord))

# gaussian spatial structure
fm_Gaus <- gls(mod_formula, 
           data = roadkills_df,
           correlation = corGaus(form=~x_coord+y_coord, nugget = T))

# linear spatial correlation 
fm_Lin <- gls(mod_formula, 
           data = roadkills_df,
           correlation = corLin(form=~x_coord+y_coord))


## compare all the spatial structures
anova(fm0, fm_CAR, fm_Gaus, fm_Lin)
# best AIC and logLik: fm_CAR
```


# 3- Run new model with spatial structure
```{r}
# previous model
mod1 <- glm.nb(nbr_carcasses ~ road_importance + distance_amenity_km + 
                 distance_water_km + woodland, 
    data = roadkills_df)
summary(mod1)
plot(mod1)
plot(ggpredict(mod1))


# new model
## without sac
model.0 <- glmmPQL(nbr_carcasses ~ road_importance + distance_amenity_km + 
                 distance_water_km + woodland,
                 random = ~1|group,
                   data=roadkills_df,
                   family = nbinom1)
summary(model.0)
plot(model.0)

## with sac
model.1 <- glmmPQL(nbr_carcasses ~ road_importance + distance_amenity_km + 
                            distance_water_km + woodland,
                   random = ~1|group,
                   data = roadkills_df,
                   correlation = corCAR1(form=~x_coord+y_coord),
                   family = nbinom1)
summary(model.1)
plot(model.1)

save(model.1, file = "data-processed/data_GLM/ngb_model_CAR.RData")

plot(Variogram(model.0), main = "No Correlation")
plot(Variogram(model.1), main = "continous autoregressive process")
```

```{r}
load("data-processed/data_GLM/ngb_model_CAR.RData")
summary(model.1)

```

